Diversity and evolution of centromere repeats in the maize genome

### 2014-07-01 response.tex

Personally I wonder whether genome size differences between maize landraces will correlate with differences in CentC content (p. 11), because it is known from old cytogenetic studies that they may vary in their number of non-centromeric heterochromatic knobs. \\

\res{We agree with the editor that the correlation between CentC abundance and knob number or genome size is of great interest. 
 Our current panel of maize lines available from the hapmap project is limited in its number of knobs, as many are inbred lines with relatively uniform knob abundance when compared to the variation seen in landraces, and therefore beleive it is not a suitable panel to investigate this question.
However, we are investigating the correlation in another project using a broad panel of approximately 300 maize landraces across an altitudinal gradient.  
Though preliminary, our as yet punpublished findings are that CentC exhibits an opposite trend to that observed in knobs; CentC increases when knob count and genome size decreases along an altitudinal gradient.  
We are currently in the process of further investigating this question, and hope to save these findings for another publication.}

\section*{Reviewer 1:}

Of particular note is the finding that "virtually all the large arrays of CentC in the maize genome derive from one of the two ancestral genomes present in modern day Zea”.  This is an interesting observation.  However, there are several potential pitfalls to this analysis that have not been addressed.  First, the authors state that "because large clusters are less likely to be misassembled, we focused our analyses on the 52 clusters $>$10KB in length."  In fact, large tandem repeat clusters are much more likely to be misassembled than shorter counterparts. The maize genome assembly was done with minimum coverage and targeted to genic regions, with gaps over repeats being ignored, meaning no attempt was made to "finish", i.e. close these gaps.  \\

@@

\res{The reviewer makes an excellent point that the most degenerate arrays are also the most likely to not be collapsed into single CentC repeats during automated assembly.
Indeed, automated assembly may explain the apparent lack of identical tandem duplicates across the genome and, more importantly, directly impact our ability to determine copy number in extremely large clusters.  
However, we would expect that any bias in assembly would impact clusters from the different subgenomes similarly and therefore should not impact our overall findings regarding subgenome.  }

\res{Furthermore, our ability to judge subgenome of origin is not likely to be impacted through assembly issues. In most cases, the origin of CentC clusters were determined with adjacent single copy markers specific to one of the subgenome as outlined in Schnable et al 2011.  The number of CentC's in the cluster would not change the origin call.}

\res{Regarding diversity of CentC copies from chromosomes with automated assembly, we also do not see striking differences.  
Centromeres 2 and 5 were carefully assembled BAC-by-BAC as part of the reference genome (see Wolfgruber et al. 2009).  CentC in these centromeres  show similar levels of diversity across CentC repeats compared to other chromosomes.  
We believe that the relatively uniform level of diversity observed across CentC is likely due to mutation over long periods of time.  
Though assembly errors likely impact our ability to detect many of the tandem duplication instances of CentC in the maize genome, we believe that identical tandem duplicates account for a relatively small quantity of overall CentC copies as evidenced by centromeres 2 and 5. }

Also, tandem repeats are often broken up by retrotransposon insertions, so that large uninterrupted CentC runs could be interpreted quite differently as "immune to retrotransposon insertion" (although this is certainly not what I suggest is going on).\\

\res{The reviewer's point addressing novel insertions breaking up large clusters is very insightful.
To investigate this, we reanalyzed cluster size data allowing for variable distances between clusters and variable cluster size to determine membership.
The results are included in a new supplementary figure, and its consequences to further analyses such as subgenome counts are discussed later in our response.}

@@

In the abstract, the authors are even bolder with the following statement: "Although the two ancestral subgenomes of maize have contributed nearly equal numbers of centromeres, our analysis shows that the vast majority of all CentC repeats derive from one of the parental genomes.”\\

\res{As we explain above, we believe our ability to determine subgenome of origin is not comprised through assembly errors, and our reanalysis considering clusters of varying size and distance supports our original idea.
The signal is likely driven by the strong bias of CentC from subgenome, showing that regardless of our definition of tandem clusters, CentC membership is biased to subgenome 1.
We have included a new supplementary table in the text to report these findings, and have altered the text to include the reanalysis.}

Wild maize relatives (other subspecies) have the same five centromeres derived from each of the two maize ancestral sub genomes as the reference inbred genome, yet the authors themselves show by FISH that ALL centromeres of these wild species have similarly high amounts of CentC.   This conclusion, therefore, is not justified in my mind.  \\

\res{We do acknowledge that FISH signal does indicate higher copy numbers of CentC in wild Zea across all chromosomes, but examination of the signals does not suggest uniformity across chromosomes.  
For example, in figure 4B, the fluorescent signal on parviglumis chromosome 2 is much stronger than chromosome 3.  
The centromeres on these chromosomes originate from different subgenomes.  
We believe a qualitative measure such as fluorescence is not a good measure for precise CentC content and are hesitant to try to compare CentC content quantitatively using FISH alone.}

Also, I agree with the authors that "Centromere inactivation could then lead to differential loss of CentC clusters", but wouldn't the same centromeres be inactivated in the wild relative as in the maize inbred? In other words, the inactive centromeres have lost CentC long before the domesticated maize diverged from the wild ancestors.  \\In sum, I find the observation of these larger CentC clusters originating from one ancestral genome extremely interesting, but am not sure it has been proven to be real (as opposed to an artifact of the genome assembly) or explained quite satisfactorily.\\

\res{
We have also modified the text regarding centromere inactivation.  
Similar numbers of currently active centromeres come from subgenome 2 (4 centromeres) and subgenome 1 (6 centromeres).  
Moreover, evidence from the fully sequenced centromere 5 suggests that large CentC arrays are often found outside of the active kinetochore.  
The bias in CentC content is thus likely to be the result of subgenome-wide patterns of deletion rather than centromere inactivation.
} %JEFFFF}

### 2014-07-01 centc.tex
Final before resubmit

We designed a simulation to estimate the accuracy of our measurements of CentC content (code available at: \url{https://github.com/kddistor/dnasims}).  
In short, our simulations altered the copy number of CentC repeats over a region of fixed length (10Mb), changing the percentage of the genome deriving from the repeat.  
Illumina reads were simulated from each of the DNA strings and mapped using our pipeline.  These simulations showed that our pipeline captured relative differences in abundance well, but underestimated total abundance of CentC.  
We found that our pipeline could accurately capture differences of 0.05\% change in CentC abundance, suggesting that larger differences are likely to be biologically real (Figure \ref{Supp_Accuracy}).

@@

\begin{figure}[h]
\centering
% For two-column wide figures use
% Use the relevant command to insert your figure file.
\includegraphics[width=1\textwidth]{Fig3_Heatmap}
\caption{CentC physical location and genetic relatedness for (a) chromosome 2 and (b) chromosome 5.  On the physical map above, red lines show locations of numbered CentC clusters and blue blocks show the location of the active kinetochores.  Scale bar is in MB.  Below each physical map is shown a heatmap of genetic relatedness of each CentC to (top row) other copies within its island of tandem repeats delineated by dotted lines and (bottom row) all other copies on the chromosome.  Darker colors indicate higher relatedness.  The total number of CentC in each cluster is shown below the map
}
\label{heatmap}    
\end{figure}

>>>>

We designed a simulation to estimate the accuracy of our measurements of CentC content (code available at: \url{https://github.com/kddistor/dnasims}).  
In short, our simulations altered the copy number of CentC repeats over a region of fixed length (10 Mb), changing the percentage of the genome deriving from the repeat.  
Illumina reads were simulated from each of the DNA strings and mapped using our pipeline.  These simulations showed that our pipeline captured relative differences in abundance well, but underestimated total abundance of CentC.  
We found that our pipeline could accurately capture differences of 0.05\% change in CentC abundance, suggesting that larger differences are likely to be biologically real (Figure \ref{Supp_Accuracy}).

@@

\begin{figure}[h]
\centering
% For two-column wide figures use
% Use the relevant command to insert your figure file.
\includegraphics[width=1\textwidth]{Fig3_Heatmap}
\caption{CentC physical location and genetic relatedness for (a) chromosome 2 and (b) chromosome 5.  On the physical map above, red lines show locations of numbered CentC clusters and blue blocks show the location of the active kinetochores.  Scale bar is in Mb.  Below each physical map is shown a heatmap of genetic relatedness of each CentC to (top row) other copies within its island of tandem repeats delineated by dotted lines and (bottom row) all other copies on the chromosome.  Darker colors indicate higher relatedness.  The total number of CentC in each cluster is shown below the map
}
\label{heatmap}    
\end{figure}
